Multiple Dice Working as One: CAD Flows and Routing Architectures for Silicon Interposer FPGAs

### 2015-03-04 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Updated hierarchy in section V
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4587 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

\subsection{CAD Flow Variations}
The overall CAD flow, including the partitioning steps, is shown in Figure \ref{fig:partitioning_cad_flow}.
\begin{figure}[!htbp]
\centering
\includegraphics[height=3.5in]{partitioning_cad_flow.eps}
\caption{Possible ways to add partitioning to the VPR CAD flow.}
\label{fig:partitioning_cad_flow}
\end{figure}

\subsubsection{Partitioner Stage}\label{sec:partitioner_stage}
As shown in Figure \ref{fig:partitioning_cad_flow}, we explore the effect of circuit partitioning at two different stages of the CAD flow. The first possibility is to partition the netlist of primitive elements (LUTs, FFs, RAM slices, etc.) before packing into the larger function blocks such as logic blocks and (32 kb) RAM blocks. The partitioning constraints must be respected by the packer; it must not pack two primitives assigned to different partitions (dice) into the same function block. We updated the VPR clustering algorithm to respect this new constraint. The second possible way to incorporate partitioning is after packing, where partitioning is applied to the function block (cluster-level) netlist. Each partition is then assigned (randomly) to a separate FPGA die. The partitioning constraints must then be respected by the placer; it is free to move function blocks around within a die, but not across interposer cutlines. These two partitioning flow options can be combined -- if partitioning is applied before packing, one can also pass the partition information to the placement stage. The post-packing placement constraints can be derived from the primitive block partitioning constraints, with the valid placement region for a given function block being equal to the intersection of the valid placement regions of the primitives packed into that function (clustered) block. There are, therefore, four possible partitioning flows:

@@

\subsubsection{Partitioning Tool}
Two well-known freely available graph partitioning tools are Metis~\cite{karypis1998multilevelmetis} and hMetis~\cite{karypis1999multilevelhmetis}. Metis operates on graphs, while hMetis operates on hypergraphs -- a hypergraph is a generalization of a graph, where each edge can connect any number of vertices. It is most natural to represent a circuit netlist as a hypergraph, where the vertices are netlist pins (or blocks) and the hyperedges are the nets joining the pins. Owing to the natural hypergraph representation, hMetis would be the most appropriate choice of partitioner. However, there are other issues to consider for our use case.

@@

\subsubsection{Hypergraph to Graph Transformation}\label{sec:hypergraph_to_graph}
We transformed the circuit netlist hypergraph to a graph in the following way:
\begin{enumerate}[label=(\roman*)]
\item For each hyperedge of the circuit netlist hypergraph, we generated a graph (the \emph{per-net subgraph}).
\item We combined all of the per-net subgraphs, summing the edge weights for edges appearing in more than one per-net subgraph, to generate the total netlist graph.
\end{enumerate}

@@

\subsubsection{Partitioner Stage Results}\label{sec:partitioner_stage_results}
We compare the performance of the four CAD flow variations described in Section \ref{sec:partitioner_stage} on the 8 largest VTR benchmarks. We again use an unbalance of $5\%$ and split the clustered netlist into two partitions (constraining the packer and/or the placer, as required by the flow under test). The interposer parameters (\% wires cut = $80\%$ and delay increase = 1 ns) and bloat factor ($10\%$ in each dimension) are the same as in Section \ref{sec:hypergraph_to_graph}. Figure \ref{fig:flows_mcw} shows the minimum routable channel width for each partitioning CAD flow variation.

@@

\subsubsection{\% wires cut}
To see how well the CAD flow works for a range of interposer architectures, we varied the fraction of wires cut and measured the average minimum channel width over the 8 largest VTR circuits. As shown in Figure \ref{fig:wires_cut}, this new flow is a significant improvement over that of~\cite{interposer2014} for FPGAs where more than half the wires are cut at interposer boundaries. It is slightly worse than the traditional FPGA CAD flow even for an architecture with 0\% of the wires cut, which is a monolithic FPGA, presumably because we have restricted the placer's freedom.

\begin{figure}[!htbp]
\centering
\input{wires_cut}
\caption{Impact of partitioning-based CAD flow on routability for a range of architectures.\protect\footnotemark}
\label{fig:wires_cut}
\end{figure}
\footnotetext{The best flow from~\cite{interposer2014} does not provide data at 90\% wires cut.}
\subsubsection{Bloating of the Clustered Netlist due to Partitioning Constraints}
The additional constraints on the packer imposed by partitioning may result in a less dense packing. Figure \ref{fig:packing_bloat} compares the number of complex blocks in the partitioning-constrained and unconstrained clustered netlists.

>>>>

The overall CAD flow, including the partitioning steps, is shown in Figure \ref{fig:partitioning_cad_flow}.
\begin{figure}[!htbp]
\centering
\includegraphics[height=3.5in]{partitioning_cad_flow.eps}
\caption{Possible ways to add partitioning to the VPR CAD flow.}
\label{fig:partitioning_cad_flow}
\end{figure}

\subsection{Partitioner Stage}\label{sec:partitioner_stage}
As shown in Figure \ref{fig:partitioning_cad_flow}, we explore the effect of circuit partitioning at two different stages of the CAD flow. The first possibility is to partition the netlist of primitive elements (LUTs, FFs, RAM slices, etc.) before packing into the larger function blocks such as logic blocks and (32 kb) RAM blocks. The partitioning constraints must be respected by the packer; it must not pack two primitives assigned to different partitions (dice) into the same function block. We updated the VPR clustering algorithm to respect this new constraint. The second possible way to incorporate partitioning is after packing, where partitioning is applied to the function block (cluster-level) netlist. Each partition is then assigned (randomly) to a separate FPGA die. The partitioning constraints must then be respected by the placer; it is free to move function blocks around within a die, but not across interposer cutlines. These two partitioning flow options can be combined -- if partitioning is applied before packing, one can also pass the partition information to the placement stage. The post-packing placement constraints can be derived from the primitive block partitioning constraints, with the valid placement region for a given function block being equal to the intersection of the valid placement regions of the primitives packed into that function (clustered) block. There are, therefore, four possible partitioning flows:

@@

\subsection{Partitioning Tool}
Two well-known freely available graph partitioning tools are Metis~\cite{karypis1998multilevelmetis} and hMetis~\cite{karypis1999multilevelhmetis}. Metis operates on graphs, while hMetis operates on hypergraphs -- a hypergraph is a generalization of a graph, where each edge can connect any number of vertices. It is most natural to represent a circuit netlist as a hypergraph, where the vertices are netlist pins (or blocks) and the hyperedges are the nets joining the pins. Owing to the natural hypergraph representation, hMetis would be the most appropriate choice of partitioner. However, there are other issues to consider for our use case.

@@

\subsection{Hypergraph to Graph Transformation}\label{sec:hypergraph_to_graph}
We transformed the circuit netlist hypergraph to a graph in the following way:
\begin{enumerate}[label=(\roman*)]
\item For each hyperedge of the circuit netlist hypergraph, we generated a graph (the \emph{per-net subgraph}).
\item We combined all of the per-net subgraphs, summing the edge weights for edges appearing in more than one per-net subgraph, to generate the total netlist graph.
\end{enumerate}

@@

\subsection{Partitioner Stage Results}\label{sec:partitioner_stage_results}
We compare the performance of the four CAD flow variations described in Section \ref{sec:partitioner_stage} on the 8 largest VTR benchmarks. We again use an unbalance of $5\%$ and split the clustered netlist into two partitions (constraining the packer and/or the placer, as required by the flow under test). The interposer parameters (\% wires cut = $80\%$ and delay increase = 1 ns) and bloat factor ($10\%$ in each dimension) are the same as in Section \ref{sec:hypergraph_to_graph}. Figure \ref{fig:flows_mcw} shows the minimum routable channel width for each partitioning CAD flow variation.

@@

\subsection{\% wires cut}
To see how well the CAD flow works for a range of interposer architectures, we varied the fraction of wires cut and measured the average minimum channel width over the 8 largest VTR circuits. As shown in Figure \ref{fig:wires_cut}, this new flow is a significant improvement over that of~\cite{interposer2014} for FPGAs where more than half the wires are cut at interposer boundaries. It is slightly worse than the traditional FPGA CAD flow even for an architecture with 0\% of the wires cut, which is a monolithic FPGA, presumably because we have restricted the placer's freedom.

\begin{figure}[!htbp]
\centering
\input{wires_cut}
\caption{Impact of partitioning-based CAD flow on routability for a range of architectures.\protect\footnotemark}
\label{fig:wires_cut}
\end{figure}
\footnotetext{The best flow from~\cite{interposer2014} does not provide data at 90\% wires cut.}
\subsection{Bloating of the Clustered Netlist due to Partitioning Constraints}
The additional constraints on the packer imposed by partitioning may result in a less dense packing. Figure \ref{fig:packing_bloat} compares the number of complex blocks in the partitioning-constrained and unconstrained clustered netlists.
